---
title: "Demonstation of 'get_mode' function"
---

R does not have a built in mode function. Being able to calculate the modal value of a variable with a function is useful, for example, when setting covariates at a constant value to calculate predicted probabilities. 

The following function finds the mode of a given vector provided by the `data` argument. 


```{r}
# Define function to take the mode of a given vector
get_mode <- function(data, return_table = FALSE){ 
    if(is.vector(data) == FALSE) { # Check that the data is a vector, and print out an error message if it is not (for example the user has entered a whole data frame)
        print("Error: the object provided in the data argument is not a vector.")
    } 
    else{
        frequency_table <- table(data) # Tabulate the frequency of each unique value in the vector
        max_frequency <- max(frequency_table) # Take the most frequent value
        mode <- names(frequency_table[frequency_table == max_frequency]) # Filter the name of the most frequent (modal) value
        if(return_table == TRUE){
          frequency_table <- data.frame(frequency_table)
          frequency_table <- frequency_table[order(frequency_table$Freq, decreasing = TRUE),]
          return(frequency_table)
        } else{
            return(mode) 
        }  
    }
}
```

The second parameter `return_table` allows the user to return the full frequncy table, instead of the modal value. The code below runs the code as default (with `return_table = FALSE`). 

```{r}
get_mode(data = quakes$station)
```

The output simply returns `14` which is the modal station number in the dataset. 

By contrast, when the `return_table` argument is set to `TRUE` the function shows the frequency of all values in the vector, ordered by frequency. 

```{r}
get_mode(data = quakes$station, return_table = TRUE)
```